---
title: "database connection"
output: html_document
---
These are the packages we will need to run sql in R and to connect to the server

```{r setup}
#library(tidyverse)
library(mosaic)
library(dplyr)
require("RPostgreSQL")
#set up the driver --depends on the database installation (mysql, access, sqlserver are other options)
driver = dbDriver("PostgreSQL")
#set up a connection by "loggin in" the to server.  The connection is your link to the database
############# GO AHEAD AND USE user="rhodes" #################
conn = dbConnect(driver,dbname="pres",host="itcsdbms.juniata.edu",user="rhodes", port=5432)

```

Simple commands.
Verify that the database tables are what are expected by printing the connection variable

```{r}
conn
#Can look at existence of any table in the database
dbExistsTable(conn,'presidents')
```

Simple selecting/printing of a database table 

We will use a sql chunk. The format inside {} is {sql, connection=ConnectionName, output.var="NameForOutput"} It is possible to do with this within R chunks using dbGetQuery().
```{sql, connection=conn, output.var="PRES"}
SELECT * FROM presidents
```

Now you can access the data within R with it's name PRES as usual.
```{r}
PRES
str(PRES)
```

Can store the results in a tidydata set

Don't confuse the SQL select --the primary retrieval command in SQL
with select function in R -- choose columns from a data set
```{sql connection=conn }
select *, pop from states
```

```{r}
tbl = dbGetQuery(conn, "select * from states")
tbl
select(tbl, state, pop)
```

```{sql, connection=conn}
SELECT state,pop FROM states
```
but you can do the column selection in the SQL select
```{r}
query = "SELECT P.Pres,spouse    
FROM presidents p   
WHERE p.birthdate>'1900-jan-1'"
stateTbl = dbGetQuery(conn, query)
stateTbl
```


1. Find names of presidents who were Democrat.
```{r}

```

2. Find names of presidents who were neither Republican or Democrat.
```{r}

```

3. Find names and spouses of presidents who were from California.
```{r}

```

4. Find home states and state populations whose presidents are still living.
```{r}

```

5. Find home states and populations of Democrat presidents.
```{r}

```

6. Find presidents whose home state's population is greater than 10,000,000.
```{r}

```

7. Find states that have less than 20 electoral votes and had a president.
```{r}

```

8. Find the years when the Republican party won the election, show the president, years and their home  states.
```{r}

```

